### 2022-02-22 manuscript.Rmd
initial commit

---
title: "comparative_RNA-Seq_manuscript"
author: "Thomas_Nicholson"
date: "19/02/2022"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggtree)
library(treeio)
library(ape)
library(UpSetR)
library(svglite)
library(ROSE)
library(ggpubr)
source('~/bin/R/functions.R')

data_path = '~/bin/R/r_files/'
figure_path = '~/Google Drive/My Drive/Comparative_RNA-Seq/'
```

## UpsetR figure

```{r upsetr_tree}
#ML phylogenetic tree built using 16s rRNA for the Gammaproteobacteria
#The full tree was also built for all bacterial genera, but is not displayed
tree <- read.tree(paste0(data_path, 'upsetr.tree'))

  tbl_tree <- as_tibble(tree)

#list of extra Gammaproteobacteria that were 
#not included in the RNA-sed analysis
to_drop <- c("Azotobacter", "Marinobacter", "Pseudoalteromonas", "Agarivorans", 
             "Vibrio", "Alishewanella", "Aggregatibacter", "Mannheimia", 
             "Actinobacillus", "Xenorhabdus", "Providencia", "Proteus",
             "Pantoea", "Brenneria", "Lonsdalea", "Buchnera.", "Wigglesworthia", 
             "Sodalis", "Dickeya", "Citrobacter", "Plautiasymbiont", 
             "Shewanella", "Moritella", "Moraxella", "Psychrobacter", 
             "Methylomonas", "Cycloclasticus", "Methylococcus", "Francisella",
             "Pseudoxanthomonas", "Candidatus", "Plautia", "Methylophaga", 
             "Pasteurella", "Salinivibrio")

  sub_tree <- drop.tip(tree, to_drop)

    tbl_sub_tree <- as_tibble(sub_tree)

p <- ggtree(sub_tree) + 
  geom_tiplab(align = T) +
  xlim(0, 0.35)

p
if(FALSE){
ggsave(filename = paste0(figure_path, "SVG/upsetr_subset_tree.svg"), 
       plot = p, width = 8, height = 16)
}
p <- ggtree(tree) + 
  geom_tiplab(align = T) +
  xlim(0, 0.5)

p
if(FALSE){
ggsave(filename = paste0(figure_path, "SVG/upsetr_tree.svg"), 
       plot = p, width = 8, height = 16)
}
```

```{r upsetr_plot}
#matrix of bool values indicating for each RNA if it was found in a genus
load(paste0(data_path, "upsetSubsetPC.Rda"))
load(paste0(data_path, "upsetSubsetPredicted.Rda"))

#list of extra Gammaproteobacteria that were 
#not included in the RNA-sed analysis
genera_arranged <- c("Lysobacter", "Stenotrophomonas", "Xylella", "Xanthomonas", 
                     "Methylomicrobium", "Pseudomonas", "Acinetobacter", 
                     "Alteromonas", "Photorhabdus", "Yersinia", "Erwinia", 
                     "Edwardsiella", "Serratia", "Klebsiella", "Enterobacter", 
                     "Salmonella", "Shigella", "Escherichia")

upset_pred <- select_columns_by_list(upsetSubsetPredicted, genera_arranged)
upset_pc <- select_columns_by_list(upsetSubsetPC, genera_arranged)

UpSetR::upset(upset_pc, sets = colnames(upset_pc), 
              mb.ratio = c(0.55, 0.45), order.by = "freq", keep.order = T)
UpSetR::upset(upset_pred, sets = colnames(upset_pred), 
              mb.ratio = c(0.55, 0.45), order.by = "freq", keep.order = T)

if(FALSE){
svglite(filename=paste0(figure_path, "SVG/upsetr_pc.svg"))
UpSetR::upset(upset_pc, sets = colnames(upset_pc), 
              mb.ratio = c(0.55, 0.45), order.by = "freq", keep.order = T)
dev.off()
svglite(filename=paste0(figure_path, "SVG/upsetr_pred.svg"))
UpSetR::upset(upset_pred, sets = colnames(upset_pred), 
              mb.ratio = c(0.55, 0.45), order.by = "freq", keep.order = T)
dev.off()
}

```

```{r expressed_regions_frequency_plot}
#matrix of bool values indicating for each RNA if it was found in a genus
load(paste0(data_path, "upsetSubsetPC.Rda"))
load(paste0(data_path, "upsetSubsetPredicted.Rda"))

#list of extra Gammaproteobacteria that were not included in the RNA-sed analyis
genera_arranged <- c("Lysobacter", "Stenotrophomonas", "Xylella", "Xanthomonas", 
                     "Methylomicrobium", "Pseudomonas", "Acinetobacter", 
                     "Alteromonas", "Photorhabdus", "Yersinia", "Erwinia", 
                     "Edwardsiella", "Serratia", "Klebsiella", "Enterobacter", 
                     "Salmonella", "Shigella", "Escherichia")

upset_pred <- select_columns_by_list(upsetSubsetPredicted, genera_arranged)
upset_pc <- select_columns_by_list(upsetSubsetPC, genera_arranged)

pc_freq <- colSums(upset_pc)
pred_freq <- colSums(upset_pred)

pc_freq <- vector_to_dataframe(pc_freq, 'genera')
pred_freq <- vector_to_dataframe(pred_freq, 'genera')

pc_freq <- pc_freq %>% dplyr::rename(freq = vec) %>% mutate(group ='pc')
pred_freq <- pred_freq %>% dplyr::rename(freq = vec) %>% mutate(group ='pred')

freq_data <-pc_freq %>% bind_rows(pred_freq)

freq_data$genera <- factor(freq_data$genera, levels = unique(freq_data$genera))

p <- ggplot() +
  geom_bar(data = freq_data, 
           aes(y = genera, x = freq, group = group, fill = group), 
           stat = 'identity', position = 'dodge')

p

if(FALSE){
  ggsave(filename = paste0(figure_path, 'SVG/rnas_frequency.svg'), 
  plot = p, width = 8, height = 16)
}

```

## Random forest interpretation figure

```{r random_forest_frequency_plots}
load("~/bin/PhD/Chapter_4/chapter_4_files/predDat.Rda")
load("~/bin/PhD/Chapter_4/chapter_4_files/validatation2.Rda")

#select the desired columns from the predicted data and validatation data
probDat <- predDat %>% 
  select(probability, ID, group, srna.counts.2) %>% 
  bind_rows(validation2 %>% select(probability, ID, group, srna.counts.2))

#not sure if setting factors will break anything so using another data frame
plotDat <- probDat
plotDat$group <- factor(plotDat$group, 
                        levels = c('Positive Control', 
                                   'Predicted', 
                                   'Negative Control'))

#plot histogram of the probabilities
p <- ggplot() +
  geom_histogram(data = plotDat, 
                 aes(x = probability, 
                     y = ..density.., 
                     group = group, 
                     fill = group), 
                 binwidth = 0.02) +
  geom_vline(xintercept = 0.17) +
  geom_vline(xintercept = 0.5) 
p

if(FALSE){
  ggsave(filename = paste0(figure_path, "SVG/histogram_probabilities.svg"), 
         plot = p, width = 178, height = 155, units = "mm")
}

#get the cumulative counts of the number of alignments as probability increases
countsCumul <- cumulativeCounts(dists = probDat, 
                                smooth = F, 
                                target_column = 'probability')

##produces plot to use for figure showing probability results
p <- ggplot() +
  geom_line(data = countsCumul, aes(x= probability, 
                                    y = cumulative_prop,
                                    group = group, 
                                    colour = group),
            size = 1.5, 
            show.legend = F) +
  scale_y_continuous(trans = 'log10')
p + theme_classic()

if(FALSE){
  ggsave(filename = paste0(figure_path, "SVG/cumulative_probabilities.svg"), 
                           plot = p, width = 178, height = 155, units = "mm")
}
```

```{r random_forest_statistics}
load("~/bin/PhD/Chapter_4/chapter_4_files/predDat.Rda")
load("~/bin/PhD/Chapter_4/chapter_4_files/validatation2.Rda")

#select the desired columns from the predicted data and validatation data
probDat <- predDat %>% 
  select(probability, ID, group, srna.counts.2) %>% 
  bind_rows(validation2 %>% 
              select(probability, ID, group, srna.counts.2))

#not sure if setting factors will break anything so using another data frame
plotDat <- probDat
plotDat$group <- factor(plotDat$group, 
                        levels = c('Positive Control', 
                                   'Predicted', 
                                   'Negative Control'))
dat <- allScores(plotDat, 0.01, 'probability')

#get the FNR and PPV values at each threshold
if(FALSE){
  for(i in seq(0,1, by=0.01)){
    scores <- scoreProbabities(plotDat, 
                               threshold = i, 
                               target_column = 'probability')
    print(paste0(i, ': ', scores$fnr, ', ', scores$ppv))
  }
}

scores <- scoreProbabities(plotDat, 
                           threshold = 0.17, 
                           target_column = 'probability')
printListSubset(scores, 
                vec = c('ppv', 'fnr', 'pred_pos', 'pred_pct'), 
                startText = 'p > 0.17', round_val = 3)

scores <- scoreProbabities(plotDat, 
                           threshold = 0.5, 
                           target_column = 'probability')
printListSubset(scores, 
                vec = c('ppv', 'fnr', 'pred_pos', 'pred_pct', 'pc_pct'), 
                startText = 'p > 0.5', round_val = 3)

scores <- scoreProbabities(plotDat, 
                           threshold = 0.81, 
                           target_column = 'probability')
printListSubset(scores, 
                vec = c('ppv', 'fnr', 'pred_pos', 'pred_pct', 'pc_pct'), 
                startText = 'p > 0.81', round_val = 3)

```

```{r statistics_plots}
load("~/bin/PhD/Chapter_4/chapter_4_files/predDat.Rda")
load("~/bin/PhD/Chapter_4/chapter_4_files/validatation2.Rda")

#select the desired columns from the predicted data and validatation data
probDat <- predDat %>% 
  select(probability, ID, group, srna.counts.2) %>% 
  bind_rows(validation2 %>% 
              select(probability, ID, group, srna.counts.2))

#not sure if setting factors will break anything so using another data frame
plotDat <- probDat
plotDat$group <- factor(plotDat$group, 
                        levels = c('Positive Control', 
                                   'Predicted', 
                                   'Negative Control'))
dat <- allScores(plotDat, 0.01, 'probability')

ggplot(data = dat) + 
  geom_line(aes(x = threshold, y = ppv), color = 'blue') + 
  geom_line(aes(x = threshold, y = fnr), color = 'red')

roc.curve(response = plotDat$group[plotDat$group != 'Predicted'], 
          predicted = plotDat$probability[plotDat$group != 'Predicted'])

ggplot(data = dat) + 
  geom_line(aes(x = fnr, y = specificity), color = 'blue') + 
  geom_line(aes(x = fpr, y = sensitivity), color = 'red')

```

```{r repeat_of_cumulative_distributions}
#is there a difference between low scoring (p < 0.17 and high scoring p > 0.5) predicted rufs?

load("~/bin/Phd/Chapter_4/chapter_4_files/predDat.Rda")
load("~/bin/Phd/Chapter_4/chapter_4_files/validatation2.Rda")
##function written using max_dist as column name so each variable needs to  be renamed to this before using cumulativeDistribution()

featuresSelected <- predDat %>% 
  mutate(group = ifelse(probability > 0.75, 'high', ifelse(probability <= 0.17, 'low', 'drop'))) %>% 
  filter(group != 'drop') %>% select(group, distance, read.max.score, 
                                     cov.min.eval, z_max, motif.max.score,
                                     alifold_cov_score, mfe.score, 
                                     gc.score) %>% 
  bind_rows(validation2 %>% filter(group == 'Positive Control') %>% 
              select(group, distance, read.max.score, 
                                     cov.min.eval, z_max, motif.max.score,
                                     alifold_cov_score, mfe.score, 
                                     gc.score))

dat <- featuresSelected %>% dplyr::rename(max_dist = distance) %>% filter(max_dist <= 0.15)
distance.p <- cumulativeDistribution(dat, alternative = "two.sided", show.legend = F)
distance.p <- distance.p +
   labs(y = "Cumulative Proportion", x = "Evolutionary distance")

##known sRNAs and predicted RUFs are only selected if there is read depths. For a fair comparison, RINCs with read depths of 0 are removed.
dat <- featuresSelected %>% select(group, read.max.score) %>% dplyr::rename(max_dist = read.max.score) %>% filter(max_dist > 0)
reads.p <- cumulativeDistribution(dat, alternative = 'two.sided', show.legend = F)
reads.p <- reads.p +
   labs(y = "Cumulative Proportion", x = "Total reads")+
  scale_x_continuous(trans = "log10")
#reads.p

dat <- featuresSelected %>% mutate(max_dist = -log(cov.min.eval))
rscape.p <- cumulativeDistribution(dat)
rscape.p <- rscape.p +
   labs(y = "Cumulative Proportion", x = "Rscape covariance score")
# rscape.p

##none of the z scores are greater than 3, so the NA value of 10 is changed to 3 (then the negaive is taken for the plot)
dat <- featuresSelected %>% mutate(max_dist = ifelse(z_max == 10, -3, -z_max)) %>% select(group, max_dist)
z.p <- cumulativeDistribution(dat, show.legend = F)
z.p <- z.p +
   labs(y = "Cumulative Proportion", x = "Alifold z-score (negative energy)")

# z.p
##selected a window where the values are easier to visualise. This has removed 16 known sRNAs and 9 predicted RUFs
dat <- featuresSelected %>% dplyr::rename(max_dist = motif.max.score) %>% filter(max_dist < 1000)
motif.p <- cumulativeDistribution(dat)
motif.p <- motif.p +
   labs(y = "Cumulative Proportion", x = "Motif score")

dat <- featuresSelected %>% mutate(max_dist = -alifold_cov_score) 
alifold.cov.p <- cumulativeDistribution(dat)
alifold.cov.p <- alifold.cov.p +
   labs(y = "Cumulative Proportion", x = "Alifold covariance score")

dat <- featuresSelected %>% mutate(max_dist = -mfe.score)
mfe.p <- cumulativeDistribution(dat, alternative = 'two.sided')
mfe.p <- mfe.p +
   labs(y = "Cumulative Proportion", x = "MFE score (negative energy)")

dat <- featuresSelected %>% arrange(gc.score) %>% mutate(gc.score = round(gc.score))

highCounts <- dat %>% filter(group == "high") %>% group_by(gc.score) %>% summarise(count = n()) %>% arrange(gc.score) %>% ungroup() %>% tidyr::complete(gc.score = seq(from = 0, to = 100, by = 1), fill = list(count = 0))
lowCounts <- dat %>% filter(group == "low") %>% group_by(gc.score) %>% summarise(count = n()) %>% arrange(gc.score) %>% ungroup() %>% tidyr::complete(gc.score = seq(from = 0, to = 100, by = 1), fill = list(count = 0))

highTotal <- dat %>% filter(group == "high") %>% nrow()
lowTotal <- dat %>% filter(group == "low") %>% nrow()

highGC <- zoo::zoo(highCounts$count)
lowGC <- zoo::zoo(lowCounts$count)

smoothHigh <- zoo::rollapply(highGC, width = 10, by = 1, FUN = mean, align = "center", partial = T) 
smoothLow <- zoo::rollapply(lowGC, width = 10, by = 1, FUN = mean, align = "center", partial = T)

smoothHigh <- as.data.frame(smoothHigh) %>% mutate(x = row_number() -1) %>% mutate(group = "high") %>% dplyr::rename(y = smoothHigh) %>% mutate(y = y/highTotal)
smoothLow <- as.data.frame(smoothLow) %>% mutate(x = row_number() -1) %>% mutate(group = "low") %>% dplyr::rename(y = smoothLow) %>% mutate(y = y/lowTotal)

pcCounts <- dat %>% filter(group == "Positive Control") %>% group_by(gc.score) %>% summarise(count = n()) %>% arrange(gc.score) %>% ungroup() %>% tidyr::complete(gc.score = seq(from = 0, to = 100, by = 1), fill = list(count = 0))
pcTotal <- dat %>% filter(group == "Positive Control") %>% nrow()
pcGC <- zoo::zoo(pcCounts$count)
smoothPC <- zoo::rollapply(pcGC, width = 10, by = 1, FUN = mean, align = "center", partial = T) 
smoothPC <- as.data.frame(smoothPC) %>% mutate(x = row_number() -1) %>% mutate(group = "Positive Control") %>% dplyr::rename(y = smoothPC) %>% mutate(y = y/pcTotal)

smoothGC <- smoothHigh %>%  bind_rows(smoothLow, smoothPC)

gc.p <- ggplot() +
  geom_path(data = smoothGC, aes(x = x, y = y, group = group, color = group), size = 1, show.legend = FALSE)  + labs(y = "Proportion", x = "GC percentage")  + theme_classic()

all.p <- ggarrange(distance.p, reads.p, mfe.p, z.p, rscape.p, alifold.cov.p,  motif.p, gc.p + rremove("x.text"),
          labels = LETTERS[1:9],
          ncol = 3, nrow = 3)

all.p

if(FALSE){
ggsave(filename = paste0(figure_path, "SVG/separated_distributions.svg"), plot = all.p, width = 450, height = 307, units = "mm")
}

```

### 2022-03-06 manuscript.Rmd
large number of changes made. Investigating why render_toc is not found

@@
 ---
-title: "comparative_RNA-Seq_manuscript"
-author: "Thomas_Nicholson"
-date: "19/02/2022"
-output: pdf_document
+title: 'Comparative RNA-Seq: Signal to Noise Documentation'
+author: "T.J Nicholson & P.P Gardner"
+date: "March 2022"
+output:
+  html_document:
+    toc: yes
+    toc_depth: 4
+    toc_float: yes
+    number_sections: yes
+  html_notebook:
+    toc: yes
+    toc_depth: 4
+  pdf_document:
+    toc: yes
+    toc_depth: '4'
 editor_options:
@@
 
+```{r toc, echo=FALSE, include=F} 
+render_toc("manuscript.Rmd")
+```
+
+# Data Selection {#data_selection}
+
+## Strains and experiments
+How to get the summary table for the genera for the RNA-seq data:
+
+-   Search the [SRA](https://www.ncbi.nlm.nih.gov/sra/) section of ncbi with 
+    the genus name and ['Organism'] tag.
+
+-   Select "Send results to Run selector"
+
+-   Select the "Metadata" button of the "Total" row and "Download" Column.
+
+-   The genome accession for each added genome was looked up on 
+    [NCBI](https://www.ncbi.nlm.nih.gov/) using the strain name in the file or 
+    by search for the given experiment in the 
+    [SRA](https://www.ncbi.nlm.nih.gov/sra/) section.
+
+-   Create a folder for the genus e.g. `mkdir Escherichia`
+
+-   Create a folder using the GenBank assembly 
+    accession e.g. `mkdir GCA_002504285.1.data/`
+
+-   Create an `experiments_list.txt` file listing all of the SRA experiments 
+    for the given genome and save the file in the `genus` folder.
+
+-   Run `callPeaksforGenome.sh -g <GCA Accession> -s` from within 
+    the `genus` folder.
+
+-   List of representative genomes selected from RefSeq95
+    -   2 genomes from each genera were selected (where possible)
+
+## Predicting sRNAs
+
+-   The `callPeaksforGenome.sh` carried out a number of steps
+    -   `fasterq-dump <SRA Accession>` downloaded each of the RNA-Seq files
+        listed in the `experiments_list.txt` file.
+    -   `fetch_genomes_from_GCA.sh` downloaded the nucleotide FASTA file and
+        the annotation file in GFF3 format.
+    -   Known ncRNAs are identified using `cmscan` with models from Rfam
+    -   `sra2plot.1.0.3.sh` took these files and produced a plot file (showing
+        read depths at each nucleotide position) for each of the experiments.
+    -   `run_rnaPeakCalling.R` took these plot files and identfied regions of
+        RNA expression tht occured in intergenic regions
+    -   `combine_gff_files.R` took the expressed regions from each of the 
+        individual files (including the annotation file) and merged regions
+        where there was an overlap.
+
+-   The resulting regions are separated into `Known sRNA` and `Putative RUFs`
+    based on whether an expressed region overlapped with a previously 
+    identified ncRNA
+
+-   The negative control group (RINCs) was generated 
+    using `get_random_srna_sequences.py`
+      -   For each putative RUF, a region of the same size was selected at 
+          random from an intergenic region in the same genome
+
+```{bash get_random_srna_sequences_loop, eval = F}
+for file in *.txt; do 
+  accession=`basename $file _new_calls.txt`; 
+  get_random_srna_sequences.py -a $accession; 
+  done
+```
+
+- The nucleotide sequences for each group were obtained from the nucleotide
+  FASTA files using the coordinates of the regions
+
+## Analysis
+
+### Homology Search
+
+-   `run_sRNA_nhmmer.sh -d <group_seqs.fna> -f <path/to/sequences/> -e fna`
+    -   The `-e` flag is for the the expected extension of the input files.
+    -   This loops through all the individual sequences in each data set and
+        carries out a similarity search against all the sequences in the data
+        set.
+    -   The matched sequences are removed from the remaining search list to 
+        prevent multiple identical alignments being produced.
+
+-   `run_sRNA_nhmmer.sh -d <path/to/genomes/> -f <path/to/initial_alignment> -o` 
+    `<path_to_output>/ -e stk -E 1e-3`
+    -   The `-E` flag sets the e-value threshold for inclusion.
+    -   This gets repeated two times to generate deep alignments for the 
+        sequences from each data set.
+
+-   A check was carried out for each of the groups (known sRNAs, predicted 
+    RUFs and RINCs) that none of the alignments shared sequences.
+      -   RINCs sharing sequences with either of the other data sets were 
+          removed.
+      -   Predicted RUFs sharing sequences with the known sRNAs were removed.
+
+```{bash, eval=F}
+group='predicted' #repeat for each group
+cd ${group}/alignments/
+> ../${group}_contig_pos.txt
+for file in *.stk;
+do 
+  ID=`basename $file .stk`
+
+  # extract the contig name and position for each sequence in each stockholm
+  #alignment file
+  grep "/" $file | grep -v "//" | cut -d ' ' -f1 | sed 's/\// /g' | 
+  sed 's/-/ /g'| cut -d '|' -f2 | sort | uniq | sed -e "s/$/ $ID/" >> 
+  ../${group}_contig_pos.txt
+done
+
+```
+
+```{r eval=F}
+library(GenomicRanges)
+ncdat <- read.table("negative_control_contig_pos.txt", sep = " ", fill = T)
+
+pcdat <- read.table("positive_control_contig_pos.txt", fill = T)
+
+preddat <- read.table("predicted_contig_pos.txt", fill = T)
+
+pcdat <- reformatContigPositionData(dat = pcdat)
+ncdat <- reformatContigPositionData(dat = ncdat)
+preddat <- reformatContigPositionData(dat = preddat)
+
+nc.pc.Dat <- getOverlapIDs(queryDat =ncdat, targetDat = pcdat)
+nc.pred.Dat <- getOverlapIDs(queryDat =ncdat, targetDat = preddat)
+pc.pred.Dat <- getOverlapIDs(queryDat =preddat, targetDat = pcdat)
+
+dat <- data.frame(ids = c(nc.pc.Dat$id1, nc.pred.Dat$id1))
+
+write.table(dat, file = "not_negative_control_ids.txt", row.names = F, col.names = F, quote = F)
+
+dat2 <- data.frame(ids =  pc.pred.Dat$id1)
+
+write.table(dat2, file = "not_predicted_ids.txt", row.names = F, col.names = F, quote = F)
+
+```
+
+-   A similar check was carried out within each group to reduce redundancy
+    -   Alignments sharing a sequence were merged into a single alignment.
+
+```{bash, eval=F}
+
+```
+
+### Evolutionary Distance
+
+-   A `cmscan` of all the genomes using the RF00177 Rfam model (bacterial 
+      ribosomal SSU) was carried out.
+-   [Phylip](https://evolution.genetics.washington.edu/phylip.html) was used to 
+    build a maximum likelihood phyogenetic tree.
+      -   The `RF00177.stk` alignment was reformatted to `.phylip` format
+      -   `esl-reformat phylip RF00177.stk > RF00177.phylip`
+      -   This was used as the input for `dnadist` with an output of 
+          `RF00177.dnadist`.
+
 ## UpsetR figure
 
-```{r upsetr_tree}
+```{r upsetr_tree, eval=F}
 #ML phylogenetic tree built using 16s rRNA for the Gammaproteobacteria
@@
 
-```{r upsetr_plot}
+```{r upsetr_plot, eval=F}
 #matrix of bool values indicating for each RNA if it was found in a genus
@@
 
-```{r expressed_regions_frequency_plot}
+```{r expressed_regions_frequency_plot, eval=F}
 #matrix of bool values indicating for each RNA if it was found in a genus
@@
 
-```{r random_forest_frequency_plots}
+  -   The output of the RF classifier for the test data and the predicted RUFs
+
+```{r random_forest_frequency_plots, }
 load("~/bin/PhD/Chapter_4/chapter_4_files/predDat.Rda")
 load("~/bin/PhD/Chapter_4/chapter_4_files/validatation2.Rda")
+#select the desired columns from the predicted data and validatation data
 
-#select the desired columns from the predicted data and validatation data
 probDat <- predDat %>% 
@@
 
-```{r random_forest_statistics}
+  -   The predicted RUFs were split into <strong>transcriptional noise</strong>, 
+      <strong>possible RNA</strong>, and <strong>putative RNA</strong>.
+  -   The categories were then compared to see if there was a difference in the
+      level of transcription between the groups.
+
+```{r transcription_levels_by_result}
+
+```
+
+```{r random_forest_statistics, eval=F}
+#TODO this is not used
 load("~/bin/PhD/Chapter_4/chapter_4_files/predDat.Rda")
@@
 
-```{r statistics_plots}
+```{r statistics_plots, eval=F}
+#TODO this is not used
 load("~/bin/PhD/Chapter_4/chapter_4_files/predDat.Rda")
@@
 
-```{r repeat_of_cumulative_distributions}
+```{r repeat_of_cumulative_distributions, eval=F}
+#TODO this is not used
 #is there a difference between low scoring (p < 0.17 and high scoring p > 0.5) predicted rufs?

### 2022-03-06 bin/callPeaksforGenome.sh

@@ -1,0 +1,213 @@
+#!/bin/bash
+
+#Take in list of SRA and a GCA
+#
+#download the genome and gff
+#
+#loop through list
+#	download SRA
+#	map reads
+#	remove CDS
+#	peak calling
+#combine gffs
+
+
+FILE_PATH=`dirname $0`
+number_of_sra="10"
+output_path="./"
+CPUS='6'
+output_log=/dev/stdout
+display_available_files="F"
+use_sra_file="T"
+while getopts "g:n:o:c:sqth" arg; do
+  case $arg in
+    g)
+      gca=$OPTARG
+      #echo $file1
+      ;;
+    n)
+      number_of_sra=$OPTARG
+      #echo $output_file
+      ;;      
+    o)
+      output_path=$OPTARG
+      #echo $output_file
+      ;;
+	c)
+      CPUS=$OPTARG
+      #echo $file1
+      ;;  
+    s)
+      use_sra_file="F"
+      ;;                
+	q)
+      output_log=$gca.log
+      #echo $file1
+      ;;
+    t)
+    display_available_files="T"
+    ;; 
+    h)
+	echo 'PredVirusHost.sh: compares proteins from a number of contigs to hmm models and then scores'
+	echo 'Version 3.0 2018'
+	echo '# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -'
+	echo 'usage: [options] <filename> <file type>'
+	echo ' '
+	echo 'Basic options:'
+	echo '  -i : <filename> the protein fasta file with protein identifiers not containing spaces'
+	exit      
+      ;;
+      
+    esac
+done    
+
+
+if [[ -z $gca ]]; then
+	echo 'Error: GCA needed. Specify with -g <gca>'
+	echo ' '
+	echo 'Use -h for more help.'
+	echo ' '
+	exit
+fi
+
+
+#mkdir -p $output_path 
+
+if [[ $display_available_files == "T" ]]; then
+	grep $gca ~/phd/RNASeq/SRA_bacteria_RNAseq.txt | grep "PAIRED" | grep "Illumina HiSeq"
+	exit
+fi
+
+
+
+if [[ $use_sra_file == 'T' ]]; then      
[…]

### 2022-03-06 bin/cmscanToGffWrapper.R

@@ -1,0 +1,73 @@
+#!/usr/bin/env Rscript
+library('getopt')
+
+
+spec = matrix(c(
+  'cmscanOutput', 'f', 1, "character",
+  'gcf', 'g', 1, "character",
+  'help' , 'h', 0, "logical",
+  'file_path', 'p', 2, "character",
+  'out_name', 'o', 2, "character"
+), byrow=TRUE, ncol=4)
+
+
+opt = getopt(spec)
+#
+# opt$cmscanOutput <- "GCA_000017745.1.tblout"
+# opt$gcf <- "GCA_000017745.1"
+# opt$file_path <- "~/phd/RNASeq/escherichia/"
+# opt$output <- "escherichia_test"
+
+if ( !is.null(opt$help) ) {
+  cat("cmscanToGffWrapper.R version 1.0\n\n")
+  cat("Use cmscanToGffWrapper.R <options> -f <cmscan ouptut file> -g <gff file>\n\n")
+  cat("Options:\n")
+  cat("  -f <cmscan ouptut file> The file that contains the cmscan output\n")
+  cat("  -g <gff file> The file that contains the gff data. Do not inclue the gff file extension\n")
+  cat("  -f <file path> The location of the other files and the output file\n")
+  cat("  -o <output file name> The name of the output file. Do not inclue the gff file extension. The default is the same as the gca input\n")
+   q(status=1)
+}
+
+if ( is.null(opt$cmscanOutput) ) {
+  cat("Error: -f <cmscan ouptut file> is required.\n")
+  q(status=1)
+}
+if ( is.null(opt$gcf) ) {
+  cat("Error: -g <gff file> is required.\n")
+  q(status=1)
+}
+
+library(tidyverse)
+library(comparativeSRA)
+
+if ( is.null(opt$file_path ) ) { opt$file_path = "." }
+if ( is.null(opt$output ) ) { opt$output = opt$gcf }
+
+rfamRes <- read.table(paste(opt$file_path, opt$cmscanOutput, sep = "/"), header = F, comment.char = "#",quote = "", fill = T)
+
+
+gff <- cmscanToGff(rfamRes = rfamRes)
+
+
+gffMain <- readLines(paste(opt$file_path, "/", opt$gcf, ".gff", sep = ""))
+gffMain <- data.frame(text = gffMain)
+genomeInfo <- as.character(gffMain[8,1])
+genomeBuild <- as.character(gffMain[4,1])
+genomeSpecies <- as.character(gffMain[9,1])
[…]

### 2022-03-06 bin/combine_gff_files.R

@@ -1,0 +1,249 @@
+#!/usr/bin/env Rscript
+suppressMessages(library('getopt'))
+suppressMessages(library('dplyr'))
+
+
+# getopts -----------------------------------------------------------------
+
+
+spec = matrix(c(
+  'sra', 'f', 1, "character",
+  'gff', 'g', 1, 'character',
+  'help' , 'h', 0, "logical",
+  'stranded' , 's', 0, "logical",
+  'quiet' , 'q', 0, "logical",
+  'file_path', 'p', 2, "character",
+  'out_name', 'o', 2, "character",
+  'random_data', 'r', 1, "character"
+), byrow=TRUE, ncol=4)
+
+
+opt = getopt(spec)
+
+if ( !is.null(opt$help) ) {
+  cat("combine_gff_files.R version 1.0\n")
+  cat(" \n")
+  cat("Use combine_gff_files.R <options> -f <files>\n")
+  cat(" \n")
+  cat("Options:\n")
+  cat("  -f <files> The gff files\n")
+  cat("  -s <stranded data> The data is stranded\n")
+  cat("  -r <random data> The file to remove CDS regions from\n")
+  cat("  -q <quiet> Do not print any updates\n")
+  cat("  -p <file path> The location of the other files and the output file\n")
+  cat("  -o <output file name> The name of the output file. Do not inclue the gff file extension. The default is the same as the sra input\n")
+  q(status=1)
+}
+
+if ( is.null(opt$sra) ) {
+  cat("Error: -f <files> is required.\n")
+  q(status=1)
+}
+
+if ( is.null(opt$out_name) ) {
+  cat("Error: -o <output file name> is required.\n")
+  q(status=1)
+}
+
+
+# packages ----------------------------------------------------------------
+
+
+suppressMessages(library(tidyverse))
+suppressMessages(library(comparativeSRA))
+
+# defining variables ------------------------------------------------------
+
+
+if ( is.null(opt$file_path ) ) { opt$file_path = "." }
+if ( is.null(opt$file_path ) ) { opt$file_path = "." }
+if ( is.null(opt$random_data ) ) { opt$random_data = "" }
+if(is.null(opt$stranded)){
+  stranded <- F
+}else{
+  stranded <- T
+}
+
+if(is.null(opt$quiet)){
+  quiet <- F
+}else{
+  quiet <- T
+}
+
+
+
+# gffDat  <- tryCatch({
[…]

### 2022-03-06 bin/fetch_genomes_from_GCA.sh

@@ -1,0 +1,137 @@
+
+#!/bin/bash
+
+##Downloads the genome sequence (and gff file if the -g flag is used) when given a 
+##GCA accession number.
+
+usage(){
+    echo "fetchGenomeGCA.sh is a script for downloading a genome (and GFF file) from a GCA accession.  
+Usage:
+ fetchGenomeGCA.sh [opts] [input]
+
+Options:
+	-h	Display this help
+
+Input	       
+	-r	Reference genome accession (required)
+	-o	Output name
+	-e Fasta file extension
+   	-g include the GFF file
+
+"
+}
+
+while getopts "r:o:e:gh" arg; do
+case $arg in
+	r) 
+	GENOME=${OPTARG};;
+	o) 
+	OUTPUT=${OPTARG};;
+	e) 
+	EXTENSION=${OPTARG};;
+	g)
+      GFF='y'
+      ;;  
+    h)
+		usage
+		exit
+      ;;    
+	\?) 
+	echo "Unknown option: -${OPTARG}" >&2; exit 1;;
+    esac
+done
+
+if [ -z ${GENOME} ]; then
+    echo "Error: No input specified." >&2
+    usage
+    exit 1
+fi
+
+if [ -z ${OUTPUT} ]; then
+
+OUTPUT=${GENOME}
+
+fi
+
+if [ -z ${EXTENSION} ]; then
+
+EXTENSION="fna"
+
+fi
+
+
+AssemblyName=$(esearch -db assembly -query ${GENOME} | efetch -format docsum | xtract -pattern DocumentSummary -element AssemblyName)
+refseqID=$(esearch -db assembly -query ${GENOME} | efetch -format docsum | xtract -pattern DocumentSummary -element RefSeq)
+
+refseq1=$(echo $refseqID | head -c 7 | tail -c 3)
+refseq2=$(echo $refseqID | head -c 10 | tail -c 3)
+refseq3=$(echo $refseqID | head -c 13 | tail -c 3)
+
+
+##get fasta file
+if [ ! -f $OUTPUT.$EXTENSION ];then
+
+fastaLink="ftp://ftp.ncbi.nlm.nih.gov/genomes/all/GCF/$refseq1/$refseq2/$refseq3/$refseqID._$AssemblyName/$refseqID._$AssemblyName._genomic.fna.gz"
+
[…]

### 2022-03-06 bin/get_random_srna_sequences.py

@@ -1,0 +1,139 @@
+#!/usr/bin/python
+
+'''
+file paths are hard coded
+'''
+
+
+import sys
+from Bio import SeqIO
+import getopt
+import os
+from BCBio import GFF
+from Bio.Seq import Seq
+from Bio.Alphabet import generic_dna
+import random
+import comparativesrna as srna
+
+
+help = '''
+    {script_name} -c com_port [-o output_file] [--loglevel level]
+
+    Reads the temperature data from a radio.  The temperature data is output in csv form.
+
+    examples:
+        Read table from radio attached to com4 and write the table to the file
+        output.csv.
+
+            {script_name} -c com4 -o output.csv
+
+        Read table from radio attached to com3 and write the table to stdout. 
+        You can use IO redirection to send the contents where every you want.
+
+            # just print to the terminal 
+            {script_name} -c com3
+
+            # redirect to another file
+            {script_name} -c com3 > somefile.csv
+
+            # filter out temperatures that are -100
+            {script_name} -c com3 | grep -v '^-100' 
+
+
+    -c com_port
+    --com_port comport
+        Name of the COM port attached to the radio
+
+    -o output_file
+    --output output_file
+        If specified write the table data to the given file.  If not specified
+        the data will be written to stdout.
+
+    --loglevel critical | error | warning | info | debug | notset
+        Control the verbosity of the script by setting the log level.  Critical
+        is the least verbose and notset is the most verbose.
+
+        The default loglevel is {default_loglevel}.
+
+        These values correspond directly to the python logging module levels. 
+        (i.e. https://docs.python.org/3/howto/logging.html#logging-levels)
+
+
+    -h 
+    --help 
+        print this message
+
+'''
+
+def usage():
+    print help
+
+def rungetopts():
+    try:
+        opts, args = getopt.getopt(sys.argv[1:], "a:sqh", ["accession", "shuffle", "quiet", "help"])
+    except getopt.GetoptError as err:
+        # print help information and exit:
+        print(err) # will print something like "option -a not recognized"
+        usage()
+        sys.exit(2)
+    accession = ""
+    shuffled = False
+    for o, a in opts:
+            if o in ("-h", "--help"):
+                usage()
+                sys.exit()
+            elif o in ("-a", "--accession"):
[…]

### 2022-03-06 bin/run_Alifold.sh

@@ -1,0 +1,113 @@
+#!/bin/bash
+
+##makes alignments and running alifoldz and r-scape
+##GCA accession number.
+
+usage(){
+    echo "sraAlignAndFold.sh is a script for making a multiple sequence alignment and
+    getting secondary structure information.  
+Usage:
+ fetchGenomeGCA.sh [opts] [input]
+
+Options:
+	-h	Display this help
+
+Input	       
+	-r	Folder location
+
+"
+}
+
+while getopts "i:o:h" arg; do
+case $arg in
+	i) 
+	FOLDER=${OPTARG};;
+	o) 
+	OUTPUT=${OPTARG};;
+    h)
+		usage
+		exit
+      ;;    
+	\?) 
+	echo "Unknown option: -${OPTARG}" >&2; exit 1;;
+    esac
+done
+
+if [ -z ${FOLDER} ]; then
+	FOLDER="./"
+fi
+
+
+
+
+mkdir -p "$FOLDER/alifold/post_script"
+
+
+
+let "fileNum = 0"
+for file in alignments_rnaalifold/*.stk
+
+
+do
+
+outname=`basename $file`
+if [ -f "$FOLDER/alifold/$outname.alifold" ]; then
+	#echo "$FOLDER/alifold/$file.alifold"
+	echo "Already exists: $file"
+	continue
+else
+	echo "Checking size: $file"
+fi
+
+
+lines=`wc -l < $file`
+if (( $lines < 1 ));then
+continue
+fi
+
+
+nseqs=`esl-alistat $file | grep "Number of sequences" | cut -d ":" -f2`
+
+
+
+length=`esl-alistat $file | grep "Alignment length:" | cut -d ":" -f2`
+largest_length=`esl-alistat $file | grep "Largest" | cut -d ":" -f2`
+
+
+if (( $length < 500 )); then
+
+diffLength=`expr $largest_length - $length`
+
+if (( $diffLength > $length ));then
+echo "Alignment is poor: $file"
+continue
+fi
+
+if (( $nseqs > 5000 )); then
+echo "Skipping $file (length: $length, nseqs: $nseqs)"
+echo "$file (length: $length, nseqs: $nseqs)" >> skipped_alignments.txt
+continue
+fi
+
+
+
+echo "Running alifoldz.pl on $file (length: $length, nseqs: $nseqs)"
+
[…]

### 2022-03-06 bin/run_R-scape.sh

@@ -1,0 +1,88 @@
+#!/bin/bash
+
+##makes alignments and running alifoldz and r-scape
+##GCA accession number.
+
+usage(){
+    echo "sraAlignAndFold.sh is a script for making a multiple sequence alignment and
+    getting secondary structure information.  
+Usage:
+ fetchGenomeGCA.sh [opts] [input]
+
+Options:
+	-h	Display this help
+
+Input	       
+	-r	Folder location
+
+"
+}
+
+while getopts "i:o:h" arg; do
+case $arg in
+	i) 
+	FOLDER=${OPTARG};;
+	o) 
+	OUTPUT=${OPTARG};;
+    h)
+		usage
+		exit
+      ;;    
+	\?) 
+	echo "Unknown option: -${OPTARG}" >&2; exit 1;;
+    esac
+done
+
+if [ -z ${FOLDER} ]; then
+	FOLDER="./"
+fi
+
+
+
+
+mkdir -p "$FOLDER/rscape_out"
+
+
+
+let "fileNum = 0"
+for file in alignments_rnaalifold/alignments_*; #change to _R for positive control
+do
+
+checkname=`basename $file .stk`
+if [ -f "./rscape_out/${checkname}_1.R2R.sto" ]; then
+echo "Already exists: $file"
+continue
+else
+
+	
+nseqs=`esl-alistat $file | grep "Number of sequences" | cut -d ":" -f2`
+
+length=`esl-alistat $file | grep "Alignment length:" | cut -d ":" -f2`
+largest_length=`esl-alistat $file | grep "Largest" | cut -d ":" -f2`
+
+
+if (( $length < 500 )); then
+
+diffLength=`expr $largest_length - $length`
+
+if (( $diffLength > $length ));then
+echo "Alignment is poor: $file"
+continue
+fi
+
+if (( $nseqs > 5000 )); then
+echo "Skipping $file (length: $length, nseqs: $nseqs)"
+echo "$file (length: $length, nseqs: $nseqs)" >> skipped_alignments.txt
+continue
+fi
+
+
+echo "Running R-scape on $file (length: $length, nseqs: $nseqs)"
+	
+time R-scape --r2rall --outdir ./rscape_out/ $file > rsacpe.out
+
+fi
+fi
+
+done
+

### 2022-03-06 bin/run_RNAAlifold.sh

@@ -1,0 +1,109 @@
+#!/bin/bash
+
+##makes alignments and running alifoldz and r-scape
+##GCA accession number.
+
+usage(){
+    echo "sraAlignAndFold.sh is a script for making a multiple sequence alignment and
+    getting secondary structure information.  
+Usage:
+ fetchGenomeGCA.sh [opts] [input]
+
+Options:
+	-h	Display this help
+
+Input	       
+	-r	Folder location
+
+"
+}
+
+while getopts "i:o:h" arg; do
+case $arg in
+	i) 
+	FOLDER=${OPTARG};;
+	o) 
+	OUTPUT=${OPTARG};;
+    h)
+		usage
+		exit
+      ;;    
+	\?) 
+	echo "Unknown option: -${OPTARG}" >&2; exit 1;;
+    esac
+done
+
+if [ -z ${FOLDER} ]; then
+	FOLDER="./"
+fi
+
+
+
+
+mkdir -p "$FOLDER/alifold/post_script"
+mkdir -p "$FOLDER/RNAAlifold"
+mkdir -p alignments_rnaalifold
+
+
+let "fileNum = 0"
+for file in alignments/*.stk
+
+do
+lines=`wc -l < $file`
+if (( $lines < 1));then
+continue
+fi
+
+outname=`basename $file`
+
+
+if [ -f "$FOLDER/RNAAlifold/$outname.rnaalifold" ]; then
+	echo "Already exists: $file"
+	continue
+fi
+
+
+
+nseqs=`esl-alistat $file | grep "Number of sequences" | cut -d ":" -f2`
+
+
+
+length=`esl-alistat $file | grep "Alignment length:" | cut -d ":" -f2`
+largest_length=`esl-alistat $file | grep "Largest" | cut -d ":" -f2`
+
+
+if (( $length < 500 )); then
+
+diffLength=`expr $largest_length - $length`
+
+if (( $diffLength > $length ));then
+echo "Alignment is poor: $file"
+continue
+fi
+
+if (( $nseqs > 5000 )); then
+echo "Skipping $file (length: $length, nseqs: $nseqs)"
+echo "$file (length: $length, nseqs: $nseqs)" >> skipped_alignments.txt
+continue
+fi
+
+
+echo "Running RNAAlifold on $file (length: $length, nseqs: $nseqs)"
+
+
+esl-reformat  clustal $file  | RNAalifold --aln-stk=${file} >> ./RNAAlifold/$outname.rnaalifold
[…]

### 2022-03-06 bin/run_RNAcode.sh

@@ -1,0 +1,92 @@
+#!/bin/bash
+
+##makes alignments and running alifoldz and r-scape
+##GCA accession number.
+
+usage(){
+    echo "sraAlignAndFold.sh is a script for making a multiple sequence alignment and
+    getting secondary structure information.  
+Usage:
+ fetchGenomeGCA.sh [opts] [input]
+
+Options:
+	-h	Display this help
+
+Input	       
+	-r	Folder location
+
+"
+}
+
+while getopts "i:o:h" arg; do
+case $arg in
+	i) 
+	FOLDER=${OPTARG};;
+	o) 
+	OUTPUT=${OPTARG};;
+    h)
+		usage
+		exit
+      ;;    
+	\?) 
+	echo "Unknown option: -${OPTARG}" >&2; exit 1;;
+    esac
+done
+
+if [ -z ${FOLDER} ]; then
+	FOLDER="./"
+fi
+
+
+
+
+mkdir -p "$FOLDER/rnacode_out"
+
+
+
+let "fileNum = 0"
+for file in alignments/*.stk; #change to _R for positive control
+do
+
+checkname=`basename $file .stk`
+if [ -f "./rnacode_out/${checkname}.rnacode" ]; then
+echo "Already exists: $file"
+continue
+else
+
+	
+nseqs=`esl-alistat $file | grep "Number of sequences" | cut -d ":" -f2`
+
+length=`esl-alistat $file | grep "Alignment length:" | cut -d ":" -f2`
+largest_length=`esl-alistat $file | grep "Largest" | cut -d ":" -f2`
+
+
+if (( $length < 500 )); then
+
+diffLength=`expr $largest_length - $length`
+
+if (( $diffLength > $length ));then
+echo "Alignment is poor: $file"
+continue
+fi
+
+if (( $nseqs > 5000 )); then
+echo "Skipping $file (length: $length, nseqs: $nseqs)"
+echo "$file (length: $length, nseqs: $nseqs)" >> skipped_alignments.txt
+continue
+fi
+
+if (( $nseqs < 3 )); then
+echo "Skipping $file (length: $length, nseqs: $nseqs)"
+#echo "$file (length: $length, nseqs: $nseqs)" >> skipped_alignments.txt
+continue
+fi
+
+echo "Running RNAcode on $file (length: $length, nseqs: $nseqs)"
+	
+esl-reformat clustal $file | RNAcode --outfile ./rnacode_out/$checkname.rnacode
+
+fi
+fi
+
+done

### 2022-03-06 bin/run_rmfamscan.sh

@@ -1,0 +1,199 @@
+#!/bin/bash
+
+##Run rmfam over all fasta files
+
+usage(){
+    echo "run_rmfam_scan.sh is a script for running rmfam_scan over a set of fasta
+    files.  
+Usage:
+ run_rmfam_scan.sh [opts] [extension]
+
+Options:
+	-h	Display this help
+
+Input
+	-e extension	       
+	-a	align
+	-o output
+	-k keep tmp files
+
+"
+}
+
+align="F"
+keep="F"
+while getopts "e:ao:hk" arg; do
+case $arg in
+	e)
+		extension=${OPTARG}
+		;;
+	a) 
+		align="T"
+		;;
+    h)
+		usage
+		exit
+      ;;    
+    k)
+		keep="T"
+      	;;   
+	\?) 
+	echo "Unknown option: -${OPTARG}" >&2; exit 1;;
+    esac
+done
+
+COUNTER=0
+mkdir -p "rmfam_cmscan"
+mkdir -p "rmfam_gff_files"
+mkdir -p "rmfam_tblout"
+
+
+let "fileNum = 0"
+if [[ $extension == "fna" ]]; then
+
+	for file in *.$extension
+	do
+		if [ -f "rmfam_tblout/$file.out.tblout" ]; then
+			echo "Already exists: $file"
+ 			continue
+		fi
+		length=`grep -v ">" $file | wc -c`
+		if (( $length < 500 )); then
+			echo "Running rmfam_scan on $file (length: $length)"
+			time rmfam_scan.pl -g -f ~/phd/RNASeq/RMfam/scripts/RMfam.cm $file -o $file.out 
+			mv *.tblout rmfam_tblout
+			mv *.cmscan rmfam_cmscan
+			mv *.gff rmfam_gff_files
+		else
+			echo "Skipping: $file"
+		fi
+	done
+else
+
+
+for file in *.$extension
+
+do
+
+
+if [ -f "rmfam_tblout/$file.out.tblout" ]; then
+	echo "Already exists: $file"
+ 	continue
+else
+echo "Checking size: $file"
+fi
+
+
+lines=`wc -l < $file`
+if (( $lines < 1));then
+continue
+fi
+
+# COUNTER=$((COUNTER+1))
+# 
+# if (( $COUNTER > 100 )); then
+# echo "waiting"
[…]

### 2022-03-06 bin/run_rnaPeakCalling.2.0.R

@@ -1,0 +1,303 @@
+#!/usr/bin/env Rscript
+
+##no idea which packages are needed
+library(tidyverse)
+library(devtools)
+library(lubridate)
+library(genoPlotR)
+library(drake)
+library(ape)
+library(Biostrings)
+library(ROSE)
+library(reshape2)
+
+suppressMessages(library(tidyverse))
+suppressMessages(library(comparativeSRA))
+genera_list <- list.files("~/phd/RNASeq/genera/")
+
+genera <- "Pseudomonas"
+
+for(genera in genera_list){
+  print(genera)
+  if(genera == "Escherichia"){
+    next
+  }
+
+  
+  accession_list <- list.files(paste("~/phd/RNASeq/genera/",  genera, "/", sep = ""), pattern = ".data$")
+  accessionsDat <- data.frame(accession_list = accession_list)
+  
+  accessionsDat <- accessionsDat %>% mutate_all(as.character)
+  accession_folder <-  accessionsDat$accession_list[1]
+  
+  
+  
+  for(accession_folder in accessionsDat$accession_list){
+    accession <- unlist(strsplit(accession_folder, "\\."))[c(1,2)]
+    accession <- paste(accession, collapse = ".")
+    files <- list.files(paste("~/phd/RNASeq/genera/",  genera, "/", accession ,".data/plot_files/", sep = ""), pattern = ".plot$")
+    files_done <- list.files(paste("~/phd/RNASeq/tmp/done/", sep = ""), pattern = ".plot$")
+    
+    list_ignored <- files_done[files_done %in% files]
+    
+    if(length(list_ignored) > 0){
+      print("Skipping files already done:")
+      print(list_ignored)
+    }
+    
+    files <- files[files %in% files_done == F]
+    
+    filesDat <- data.frame(files = files)
+    
+    filesDat <- filesDat %>% filter(grepl(pattern = "ncRNA", x = files)==F,
+                                    grepl(pattern = "fwd", x = files)==F,
+                                    grepl(pattern = "rev", x = files)==F,
+                                    grepl(pattern = accession, x = files)==F) %>% mutate_all(as.character)
+    
+    
+    
+    
+    if(nrow(filesDat) == 0){
+      next
+    }
+    print(accession)
+    
+    gffDat <- read.table(paste("~/phd/RNASeq/genera/",  genera, "/", accession ,".data/gff_files/", accession, ".gff", sep = ""), sep = "\t", fill = T, comment.char = "#", quote = "")
[…]

### 2022-03-06 bin/run_sRNA_nhmmer.sh

@@ -1,0 +1,155 @@
+#!/bin/bash
+
+# Setup Variables #
+
+usage(){
+    echo "run_sRNA_nhmmer.sh is a script for running nhmmer and sorting the results.  
+Usage:
+ run_sRNA_nhmmer.sh [opts] [input]
+
+Required:	       
+	-d	<database> The nucleotide database to be searched against as the nhmmer target
+	-f	<folder> The folder containing the query file for nhmmer
+	-e	<extension> The file extension of the query file in <folder>
+
+Options:
+	-h	Display this help
+	-c	Check if the sequences are >50 nt (only works on fasta files)
+"
+}
+
+check_seq_lengths=""
+database=""
+folder=""
+extension=""
+outfolder="."
+missing=""
+exitTrue="F"
+evalue="1e-5"
+
+# User Input Options #
+
+while getopts "d:o:f:E:e:ch" arg; do
+case $arg in
+	d) 
+	database=${OPTARG};;
+	f)
+	folder=${OPTARG};;
+	o)
+	outfolder=${OPTARG};;
+	E)
+	evalue=${OPTARG};;
+	e)
+	extension=${OPTARG};;	
+	c)
+	check_seq_lengths="T";;
+    h)
+		usage
+		exit
+      ;;    
+	\?) 
+	echo "Unknown option: -${OPTARG}" >&2; exit 1;;
+    esac
+done
+
+# Tests for inputs #
+
+if [[ $database == "" ]]; then
+	missing="$missing -d <database> "
+	exitTrue="T"
+fi
+
+if [[ $folder == "" ]]; then
+	missing="$missing -f <folder> "
+	exitTrue="T"	
+fi
+
+if [[ $extension == "" ]]; then
+	missing="$missing -e <extension> "
+	exitTrue="T"
+fi
+
+if [[ $exitTrue == "T" ]];then
+	echo "$missing not found. Use -h for more help."
+	exit
+fi
+
+if [[ $extension == "stk" ]]; then
+	check_seq_lengths=""
+fi
+
+echo "extension value: $extension"
+
+if [[ $check_seq_lengths == "T" ]]; then
+
+	echo "checking for short seqs"
+	mkdir -p $folder/short_seqs
+	
+	let "fileNum = 0"
+	for file in $folder/*.$extension
+	do
+	
+	seqlength=`grep -v ^">" $file | wc -c`
+	if (( $seqlength < 50 )); then
[…]

### 2022-03-06 bin/single_fasta_files.py

@@ -1,0 +1,115 @@
+#!/usr/bin/python
+import sys
+from Bio import SeqIO
+import getopt
+import os
+from BCBio import GFF
+from Bio.Seq import Seq
+from Bio.Alphabet import generic_dna
+import random
+import pandas as pd
+import comparativesrna as srna
+
+help = '''
+    {script_name} -c com_port [-o output_file] [--loglevel level]
+
+    Reads the temperature data from a radio.  The temperature data is output in csv form.
+
+    examples:
+        Read table from radio attached to com4 and write the table to the file
+        output.csv.
+
+            {script_name} -c com4 -o output.csv
+
+        Read table from radio attached to com3 and write the table to stdout. 
+        You can use IO redirection to send the contents where every you want.
+
+            # just print to the terminal 
+            {script_name} -c com3
+
+            # redirect to another file
+            {script_name} -c com3 > somefile.csv
+
+            # filter out temperatures that are -100
+            {script_name} -c com3 | grep -v '^-100' 
+
+
+    -c com_port
+    --com_port comport
+        Name of the COM port attached to the radio
+
+    -o output_file
+    --output output_file
+        If specified write the table data to the given file.  If not specified
+        the data will be written to stdout.
+
+    --loglevel critical | error | warning | info | debug | notset
+        Control the verbosity of the script by setting the log level.  Critical
+        is the least verbose and notset is the most verbose.
+
+        The default loglevel is {default_loglevel}.
+
+        These values correspond directly to the python logging module levels. 
+        (i.e. https://docs.python.org/3/howto/logging.html#logging-levels)
+
+
+    -h 
+    --help 
+        print this message
+
+'''
+
+def usage():
+    print help
+
+def rungetopts():
+    try:
+        opts, args = getopt.getopt(sys.argv[1:], "i:f:qh", ["infile", "folder", "quiet", "help"])
+    except getopt.GetoptError as err:
+        # print help information and exit:
+        print(err) # will print something like "option -a not recognized"
+        usage()
+        sys.exit(2)
+    file = ""
+    for o, a in opts:
+            if o in ("-h", "--help"):
+                usage()
+                sys.exit()
+            elif o in ("-i", "--infile"):
+                file = a
+            elif o in ("-f", "--folder"):
+                folder = a
+            else:
[…]

### 2022-03-06 bin/sra2plot.1.0.3.sh

@@ -1,0 +1,139 @@
+#!/bin/sh
+#Downloads fastq files from SRA, trims, maps and generates plotfiles for visualisation in artemis
+
+#Dependencies:
+#curl
+#sratoolkit
+#samtools 1.6 (older versions may not work for generating plotfiles)
+#bowtie2
+#trimmomatic 0.36
+
+usage(){
+    echo "sra2plot.sh is a wrapper script for downloading, mapping and visualising RNA-seq data from the NCBI Sequence Read Archive (SRA). Currently assumes paired end reads with TruSeq3 adaptors. Path for trimmomatic needs to be set to run. 
+Usage sra2plot [opts] [input]
+    
+    Options:
+		-h	Display this help
+
+		Input	       
+	       	-r	Reference genome accession (required)
+		-s	SRA run accession or name of split fastq files (Required. Format: FILE_1.fastq FILE_2.fastq)
+    		-n	Number of cores
+
+		Turn off defaults
+		-d	Turn off download. Default: download genome and SRA from NCBI if not found in working directory. 
+			(Genome accession must be in Genbank nucleotide format: https://www.ncbi.nlm.nih.gov/Sequin/acc.html)
+		-t	Turn off trimming
+		-m	Turn off mapping
+		-p	Don't make plotfiles
+		-x	Don't cleanup files"
+}
+TPATH="/Users/thomasnicholson/bin/Trimmomatic_binary-0.36"
+OUTDIR=""
+SRA=""
+GENOME=""
+TRIM=true
+MAP=true
+PLOT=true
+CLEAN=true
+DOWNLOAD=true
+THREADS=1
+
+while getopts :s:r:n:thdmpx opt; do
+    case "${opt}" in
+	h) usage;exit;;
+	t) TRIM=false;;
+	s) SRA=${OPTARG};;
+	r) GENOME=${OPTARG};;
+	d) DOWNLOAD=false;;
+	m) MAP=false;;
+	p) PLOT=false;;
+	x) CLEAN=false;;
+	n) THREADS=${OPTARG};;
+	\?) echo "Unknown option: -${OPTARG}" >&2; exit 1;;
+	:) echo "Missing option argument for -${OPTARG}" >&2; exit 1;;
[…]

### 2022-03-07 bin/run_RNAcode.sh
Update to manuscript.Rmd,
The methods section of manuscript.Rmd now need to have the final
step of running the tools added to it, and the results section needs to
be added.

A data folder has been added, and all the data files need to be
transfered over and the paths renamed.

@@ -5,30 +5,27 @@
 
 usage(){
     echo "sraAlignAndFold.sh is a script for making a multiple sequence alignment and
-    getting secondary structure information.  
+    getting secondary structure information.
 Usage:
- fetchGenomeGCA.sh [opts] [input]
+ fetchGenomeGCA.sh -i <input_folder>
 
 Options:
+-i  Input folder where the alignment are in stk format (default is the working
+directory)
 	-h	Display this help
-
-Input	       
-	-r	Folder location
 
 "
 }
 
 while getopts "i:o:h" arg; do
 case $arg in
-	i) 
+	i)
 	FOLDER=${OPTARG};;
-	o) 
-	OUTPUT=${OPTARG};;
     h)
 		usage
 		exit
-      ;;    
-	\?) 
+      ;;
+	\?)
 	echo "Unknown option: -${OPTARG}" >&2; exit 1;;
     esac
 done
@@ -37,56 +34,37 @@ if [ -z ${FOLDER} ]; then
 	FOLDER="./"
 fi
 
-
-
-
 mkdir -p "$FOLDER/rnacode_out"
 
-
-
 let "fileNum = 0"
-for file in alignments/*.stk; #change to _R for positive control
+for file in alignments/*.stk;
 do
 
+#Check if the file has already been created. Used for cases where the
+#program crashed part way through
 checkname=`basename $file .stk`
 if [ -f "./rnacode_out/${checkname}.rnacode" ]; then
-echo "Already exists: $file"
-continue
+    echo "Already exists: $file"
+    continue
 else
+    #get number of sequences and length of sequences
+    nseqs=`esl-alistat $file | grep "Number of sequences" | cut -d ":" -f2`
 
-	
-nseqs=`esl-alistat $file | grep "Number of sequences" | cut -d ":" -f2`
+    length=`esl-alistat $file | grep "Alignment length:" | cut -d ":" -f2`
+    largest_length=`esl-alistat $file | grep "Largest" | cut -d ":" -f2`
 
-length=`esl-alistat $file | grep "Alignment length:" | cut -d ":" -f2`
-largest_length=`esl-alistat $file | grep "Largest" | cut -d ":" -f2`
+    #check the sequences are <500 nt long
+    if (( $length < 500 )); then
+        diffLength=`expr $largest_length - $length`
+        #check that the alignment is not poor
+        if (( $diffLength > $length ));then
[…]

### 2022-03-07 manuscript.Rmd
Update to manuscript.Rmd
The methods section has been mostly finished. The running of each
inidivual tool, and the results need to be added in.

A data folder is added, and the data needs to be moved over and the
paths changed.

@@
 output:
+  pdf_document:
+    toc: yes
+    toc_depth: '4'
+  html_notebook:
+    toc: yes
+    toc_depth: 4
   html_document:
@@
     number_sections: yes
-  html_notebook:
-    toc: yes
-    toc_depth: 4
-  pdf_document:
-    toc: yes
-    toc_depth: '4'
 editor_options:
@@
 library(ggpubr)
+library(rstatix)
+options(dplyr.summarise.inform = FALSE)
 source('~/bin/R/functions.R')
@@
 figure_path = '~/Google Drive/My Drive/Comparative_RNA-Seq/'
-```
-
-```{r toc, echo=FALSE, include=F} 
-render_toc("manuscript.Rmd")
 ```
@@
 
+```{r redundancy_1_R, eval=F}
+#read in the data
+dat <- read.table("predicted_contig_pos.txt", sep = " ", fill = T, as.is = T)
+dat <- dat %>% select(V1, V4)
+#split out the columns of interest
+dat <- dat %>% separate(col = V1, into = c("contig", "coordinates"), sep = "\\/", remove = F)
+dat <- dat %>% separate(col = coordinates, into = c("start", "stop"), sep = "-", remove = F)
+dat <- dat %>% dplyr::rename(srna = V4) %>% select(contig, srna, start, stop) %>% mutate(strand = "+")
+colnames(dat) <- c("contig","srna", "start", "stop", "strand")
+
+#rearrange the start and stop so th at that the stop > start
+dat <- dat %>% mutate(tmpstart = ifelse(start < stop, start, stop),
+                      tmpend = ifelse(start > stop, start, stop))
+
+#Set up the data to be used in genomic ranges overlap search
+query <- GRanges(seqnames = dat$contig,
+                 ranges = IRanges(start = dat$tmpstart, end = dat$tmpend),
+                  strand = dat$strand, query_name = dat$srna)
+lookup1 <- data.frame(id1 = dat$srna, queryHits = c(1:length(dat$srna)))
+lookup2 <- data.frame(id2 = dat$srna, subjectHits = c(1:length(dat$srna)))
+
+#Run search for overlaps and merge with original input to obtain pairs of 
+#overlaps
+tmp <- GenomicRanges::findOverlaps(query, query, type = 'any')
+tmp <- as.data.frame(tmp)
+tmp <- tmp %>% left_join(lookup1) %>% left_join(lookup2)
+tmp <- tmp %>% filter(id1 != id2)
+smallDat <- tmp %>% select(id1, id2) %>% unique()
+
+#Swap columns and combine so that every combination is present in both columns
+s2 <- smallDat
+s2$id1 <- smallDat$id2
+s2$id2 <- smallDat$id1
+smallDat <- smallDat %>% bind_rows(s2) %>% unique()
+
+ids <- unique(dat$srna)
+
+#Loop over the ids and run a check to see if there are any overlapping ids
+#to speed this up check if an id has already been seen and skip it.
+#Write a list of overlapping ids to a unqiue file
+checked_ids <- c()
+for(item in ids){
+  if(item %in% checked_ids){
+    next
+  }
+  current_ids <- groupOverlapItems(smallDat, item, current_ids = c())
+  checked_ids <- c(checked_ids, current_ids)
+    write.table(x = current_ids, 
+                file = paste("combined_alignment_ids/", 
+                             current_ids[1], 
+                             "_combined_list.txt", sep = ""), 
+                append = T, quote = F, row.names = F, col.names = F)
+}
+
+```
+
 ```{bash, eval=F}
+#move into the folder where the lists of overlapping ids are kept
+cd combined_alignments_ids/
+mkdir -p fasta
+mkdir -p hmm
+mkdir -p alignments
 
+#loop through the files that contain overlapping ids
+for file in *.txt;
+do
+max_num="0"
+outname=`basename $file _combined_list.txt`
+#Check if there is only one ID listed in a file. If so there is no merge
+#required.
+line_count=`wc -l $file | cut -d ' ' -f1`
+if (( $line_count == 1 ));then
+cp ../large_alignments/alignments/$outname.stk ./alignments
+continue
+fi
+
+#loop through the file and extract the sequences from each of the stk files
+#that correspond to the ids that are overlapping
+> fasta/$outname.fa
+while read line; 
+do 
+esl-reformat fasta ../large_alignments/alignments/$line.stk >> fasta/$outname.fa
+current_num=`esl-alistat ../large_alignments/alignments/$line.stk | grep "Number of sequences" | cut -d ' ' -f 4`
+
+#check if the current alignment is the longest
+if(( $current_num > $max_num ));
+then
+max_seq="$line.stk"
+max_num="$current_num"
+fi
+done < $file
+
+#build an hmm using the longest alignment
+hmmbuild hmm/$outname.hmm ../large_alignments/alignments/$max_seq
+
+#build an alignment using all the sequences from the overlapping ids
+hmmalign --informat fasta hmm/$outname.hmm fasta/$outname.fa | esl-alimask -g --gapthresh 0.8 -p --pfract 0.5 --pthresh 0.5 - | esl-alimanip   --lnfract 0.6 --lxfract 1.4 --lmin 50 --lmax 500 --detrunc 50 - > alignments/$outname.stk
+done
+```
+
+### File setup
+
+-   There are some files that are consistently used throughout the analysis
+  -   `genome_contig_pairs.txt` is a file that maps the genome `GCA_` ids to the 
+      contig `NC` ids.
+    -   this is necessary due to the alignment reporting contig names, but the 
+        predicted RUFs belonging to genomes rather than contigs.
+  -   `query_target_pairs.txt` consists of each alignment ID and the contig IDs
+      of each sequence within the alignment.
+
+```{bash,  eval=F}
+#loop through all the genomes in the analysis
+cd representative_genomes/
+> ../genome_contig_pairs.txt
+for file in GC*.fna;  
+do   
+  ID=`echo $file | cut -d '.' -f1,2 | cut -d "_" -f1,2`;   
+  grep ^">" $file | cut -d ' ' -f1 | sort | uniq | sed 's/>//g' | sed -e "s/$/   $ID/" >> ../genome_contig_pairs.txt;   
+done
+```
+
+```{bash query_target_pairs, eval=F}
+#for each group loop through all the alignments and extract the contig IDs
+#of all the sequences within each alignment
+group="positive_control"
+cd ${group}/alignments
+> ../query_target_pairs.txt
+for file in *.stk; 
+do 
+  ID=`echo $file | cut -d '.' -f1,2 | cut -d "_" -f1,2`; 
+  ID_2=`echo $file | cut -d '.' -f1,2 `;  
+  grep ^"#=GS" $file | sort | uniq | cut -d "/" -f1 | cut -d ' ' -f2 | rev | cut -d '|' -f1 | rev | sed -e "s/$/   $ID   $ID_2/" >> ../query_target_pairs.txt;  
+done
 ```
@@
           `RF00177.dnadist`.
+
+-   The file was not formatted in a usable way for reading into R
+      -   `cat  RF00177.dnadist | tr '\n' ' ' | sed 's/ N/\nN/g' > RF00177.dists`
+
+```{r, eval=FALSE}
+#open the dnadist file
+dat <- read.table("RF00177.dists", sep = "",
+                  header = F, fill = T, stringsAsFactors = F, as.is = T)
+dat <- dat %>% filter(!is.na(V2))
+colnames(dat)[1] <- "phylip_id"
+colnames(dat) <- c("names", dat$phylip_id)
+
+#reformat the data to use in subsequent steps
+meltDat <- melt(dat)
+colnames(meltDat) <- c("phylip_id", "query.name", "distance")
+
+#open list of phylip ids (truncated) and contig ids
+load(file = "current_ids.Rda")
+current_ids <- current_ids %>% unique()
+
+#open list of genome and contig ids
+contig_labels <- read.table("genome_contig_pairs.txt")
+colnames(contig_labels) <- c("id", "target.genome")
+
+#combine the phylip ids with the genome ids
+contig_labels <- contig_labels %>% 
+  left_join(current_ids, by = "id") %>% 
+  select(-b)
+
+#combine the genome ids with the distances data for the first column
+colnames(contig_labels) <- c("target.id", "target.genome", "phylip_id")
+meltDat <- meltDat %>% left_join(contig_labels, by = 'phylip_id')
+
+#combine the genome ids with the distances data for the second column
+colnames(meltDat)[1:2] <- c("target.name", "phylip_id")
+colnames(contig_labels) <- c("query.id", "query.genome", "phylip_id")
+meltDat <- meltDat %>% left_join(contig_labels, by = 'phylip_id')
+colnames(meltDat)[2] <- c("query.name")
+
+#take the pairs of genomes and arrange based on distances (to account for
+#multiple contigs). Sort these and select a single distance for each pair. 
+meltDat <- meltDat %>%  group_by(target.genome, query.genome) %>% arrange(as.numeric(distance)) %>% mutate(row_num = row_number()) %>% filter(row_num == 1) %>% select(-row_num)
+save(meltDat, file = "~/bin/R/r_files/distanceMelt.Rda")
+
+#Reshape the data back to a matrix (not sure this was needed but helped for 
+#visualising the problem)
+mat <- reshape2::acast(data = meltDat %>% 
+                         select(target.genome, query.genome, distance), 
+                       formula = target.genome ~ query.genome)
+distanceMat <- as.data.frame(mat)
+save(distanceMat, file = "~/phd/RNASeq/distanceMat.Rda")
+
+#include a save a load point as the previous steps were time consuming and
+#this allows for troubleshooting the next steps
+load("~/phd/RNASeq/distanceMat.Rda")
+
+#get the maximum distances between pairs of sequences for each alignment.
+max_dists_pred <- calculateMaximumDistances('predicted')
+save(max_dists_pred, file = "~/bin/R/r_files/max_dists_pred.Rda")
+max_dists_pc <- calculateMaximumDistances('pc')
+save(max_dists_pc, file = "~/bin/R/r_files/max_dists_pc.Rda")
+max_dists_nc <- calculateMaximumDistances('negative_control')
+save(max_dists_nc, file = "~/bin/R/r_files/max_dists_nc.Rda")
+
+```
 
@@
 ```{r random_forest_frequency_plots, }
-load("~/bin/PhD/Chapter_4/chapter_4_files/predDat.Rda")
+load("~/bin/manuscript/data/predDat.Rda")
 load("~/bin/PhD/Chapter_4/chapter_4_files/validatation2.Rda")
@@
 ```{r transcription_levels_by_result}
+load("~/bin/manuscript/data/predDat.Rda")
 
+#create groups withing the predDat based on the probability thresholds
+predDat <- predDat %>% 
+  mutate(group = ifelse(probability <0.17, 
+                        'transcriptional noise',
+                        ifelse(probability <= 0.81, 
+                               'possible RNA', 
+                               'putative RNA')))
+predDat$group = factor(predDat$group, levels = c("transcriptional noise",
+                                                 "possible RNA", 
+                                                 "putative RNA"))
+#boxplot of transcription levels for each group
+p <- ggplot(data = predDat) +
+  geom_boxplot(aes(y = read.max.score, group = group, fill = group),
+               outlier.alpha = 0.1) +
+  scale_y_continuous(trans = 'log10')
+p
+if(FALSE){
+  ggsave(filename = paste0(figure_path, "SVG/trascription-comparison.svg"),
+         plot = p, width = 450, height = 307, units = "mm")
+}
+#run wilcox test comparing the transcriptional noise and the other groups
+wilcox.test(x = predDat$read.max.score[predDat$group == "transcriptional noise"],
+            y = predDat$read.max.score[predDat$group == "putative RNA"])
+wilcox.test(x = predDat$read.max.score[predDat$group == "transcriptional noise"],
+            y = predDat$read.max.score[predDat$group == "possible RNA"])
 ```
@@
 #TODO this is not used
-load("~/bin/PhD/Chapter_4/chapter_4_files/predDat.Rda")
+load("~/bin/manuscript/data/predDat.Rda")
 load("~/bin/PhD/Chapter_4/chapter_4_files/validatation2.Rda")
@@
 
-load("~/bin/Phd/Chapter_4/chapter_4_files/predDat.Rda")
+load("~/bin/manuscript/data/predDat.Rda")
 load("~/bin/Phd/Chapter_4/chapter_4_files/validatation2.Rda")

### 2022-03-14 manuscript.Rmd
Spearman correlation and distance filtering added.
The section to produce the spearman correlation and heatmap were added.
The removal of alignments with a maximum distance > 0.15 and the data
used to determine the appropriate threshold are added.

Extra files (old versions of the rendered markdown) have been removed.

@@
 library(rstatix)
+library(gplots)
 options(dplyr.summarise.inform = FALSE)
@@
 
+```
+
+Data sets were filtered by distance < 0.15.
+
+-   Threshold was selected based on the distance between genomes within a family
+      -   Filtered at this stage to reduce computational time
+
+```{r, phylogenetic distance, eval=T}
+load("~/bin/manuscript/data/allDists.Rda")
+maxsGenus <- allDists %>% group_by(Genus.x, Genus.y) %>% 
+  summarise(max_dist = max(distance)) %>% 
+  mutate(same.taxa = ifelse(Genus.x == Genus.y, T, F))
+maxsFamily <- allDists %>% group_by(Family.x, Family.y) %>% 
+  summarise(max_dist = max(distance)) %>% 
+  mutate(same.taxa = ifelse(Family.x == Family.y, T, F))
+maxsOrder <- allDists %>% group_by(Order.x, Order.y) %>% 
+  summarise(max_dist = max(distance)) %>% 
+  mutate(same.taxa = ifelse(Order.x == Order.y, T, F))
+maxsClass <- allDists %>% group_by(Class.x, Class.y) %>% 
+  summarise(max_dist = max(distance)) %>% 
+  mutate(same.taxa = ifelse(Class.x == Class.y, T, F))
+maxsPhylum <- allDists %>% group_by(Phylum.x, Phylum.y) %>% 
+  summarise(max_dist = max(distance)) %>% 
+  mutate(same.taxa = ifelse(Phylum.x == Phylum.y, T, F))
+maxsKingdom <- allDists %>% group_by(Kingdom.x, Kingdom.y) %>% 
+  summarise(max_dist = max(distance)) %>% 
+  mutate(same.taxa = ifelse(Kingdom.x == Kingdom.y, T, F))
+genus1 <- maxsGenus %>% ungroup() %>% filter(same.taxa, max_dist > 0) %>% select(max_dist) %>% mutate(group = "Genus")
+family1 <- maxsFamily %>% ungroup() %>% filter(same.taxa, max_dist > 0) %>% select(max_dist) %>% mutate(group = "Family")
+order1 <- maxsOrder %>% ungroup() %>% filter(same.taxa, max_dist > 0) %>% select(max_dist) %>% mutate(group = "Order")
+class1 <- maxsClass %>% ungroup() %>% filter(same.taxa, max_dist > 0) %>% select(max_dist) %>% mutate(group = "Class")
+phylum1 <- maxsPhylum %>% ungroup() %>% filter(same.taxa, max_dist > 0) %>% select(max_dist) %>% mutate(group = "Phylum")
+kingdom1 <- maxsKingdom %>% ungroup() %>% filter(same.taxa, max_dist > 0) %>% select(max_dist) %>% mutate(group = "Kingdom")
+
+maxs1 <- genus1 %>% bind_rows(family1, order1, class1, phylum1, kingdom1)
+maxs1$group <- factor(maxs1$group, levels = c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus"))
+
+p <- ggplot() +
+geom_boxplot(data = maxs1 , aes(x = group, y = max_dist), fill = NA) + coord_flip() +
+  geom_hline(yintercept = 0.15)
+p
+```
+
+### Correlation heatmap
+
+A spearman correlation matrix was produced comparing multiple metrics for each 
+feature. The best performing metrics (measured by comparing correlation with
+known sRNAs were kept).
+
+```{r correlation_heat_map, eval=T}
+load("~/bin/PhD/Chapter_4/chapter_4_files/randomForestDat.Rda")
+
+dat <- dat %>% select(-ID, -power, -motif_count, -read.counts) %>% unique() %>% 
+  mutate(cov.min.eval = -log(cov.min.eval), 
+         z_mean = -z_mean,
+         z_max = -z_max,
+         alifold.score  =-alifold.score,
+         mfe.score = -mfe.score,
+         alifold_cov_score = -alifold_cov_score)
+
+set.seed(101)
+matNames <- c("MFE score", "G+C Percentage", "Evolutionary Distance", 
+              "Read depth (max)", "Read depth (mean)",
+                        "R-scape covariance (mean score)", 
+              "R-scape covariance (max score)", 
+              "R-scape covariance (number of significant pairs)", 
+              "Motif score (mean)", "Motif score (max)", 
+              "Alifold Z score (mean)", "Alifold Z score (max)", 
+              "Alifold covariance score", "Alifold score", 
+              "Number of sequences in alignment", "Sum of motif scores", 
+              "Random", "Known sRNA or RINC")
+
+matrixList <- calculateCorrelationMatrix(dat, matNames)
+rhoMatrix <- matrixList$rhoMatrix
+sigMatrix <- matrixList$sigMatrix
+
+heatmap.2(rhoMatrix, cellnote=round(rhoMatrix, digits = 2),
+          notecex=1.5,notecol="black", col=rev(redblue(40)), 
+          density.info="none", trace="none", dendrogram=c("column"), 
+          symm=F,symkey=T,symbreaks=T, scale="none", key.title = "", 
+          srtRow=45, adjRow=c(0, 1), srtCol=45, adjCol=c(1,1), 
+          breaks=(-20:20)/20, margins = c(8, 8), cexRow=1.5, 
+          cexCol=1.5,font=1)
+
+if(FALSE){
+svglite::svglite(paste0(figure_path, "/heatmap.svg"))
+heatmap.2(rhoMatrix, cellnote=sigMatrix,notecex=1.5,notecol="black",
+          col=rev(redblue(40)), density.info="none", trace="none", 
+          dendrogram=c("column"), symm=F,symkey=T,symbreaks=T, scale="none", 
+          key.title = "", srtRow=45, adjRow=c(0, 1), srtCol=45, adjCol=c(1,1), 
+          breaks=(-20:20)/20, margins = c(8, 8), cexRow=1.5, cexCol=1.5,font=2)
+
+dev.off()
+}
+```
+
+```{r subset_correlation_heat_map, eval=T}
+load("~/bin/PhD/Chapter_4/chapter_4_files/randomForestDat.Rda")
+
+dat <- dat %>% select(group, 
+                      read.max.score, 
+                      distance, 
+                      alifold_cov_score,
+                      cov.min.eval,
+                      motif.sum,
+                      z_max,
+                      mfe.score,
+                      gc.score) %>% 
+  unique() %>% 
+  mutate(cov.min.eval = -log(cov.min.eval), 
+         z_max = -z_max,
+         mfe.score = -mfe.score,
+         alifold_cov_score = -alifold_cov_score)
+
+set.seed(101)
+matNames <- c("Read depth (max)", "Evolutionary distance", 
+              "Alifold covariance score", "R-scape covariance score", 
+              "Sum of motif scores", "Alifold Z score (max)", "MFE score", 
+              "G+C Percentage", "Random", "Known sRNA or RINC")
+
+matrixList <- calculateCorrelationMatrix(dat, matNames)
+rhoMatrix <- matrixList$rhoMatrix
+sigMatrix <- matrixList$sigMatrix
+
+heatmap.2(rhoMatrix, cellnote=round(rhoMatrix, digits = 2),
+          notecex=1.5,notecol="black", col=rev(redblue(40)), 
+          density.info="none", trace="none", dendrogram=c("column"), 
+          symm=F,symkey=T,symbreaks=T, scale="none", key.title = "", 
+          srtRow=45, adjRow=c(0, 1), srtCol=45, adjCol=c(1,1), 
+          breaks=(-20:20)/20, margins = c(8, 8), cexRow=1.5, 
+          cexCol=1.5,font=1)
+
+if(FALSE){
+svglite::svglite(paste0(figure_path, "/svg/heatmap_subset.svg"))
+heatmap.2(rhoMatrix, cellnote=sigMatrix,notecex=1.5,notecol="black",
+          col=rev(redblue(40)), density.info="none", trace="none", 
+          dendrogram=c("column"), symm=F,symkey=T,symbreaks=T, scale="none", 
+          key.title = "", srtRow=45, adjRow=c(0, 1), srtCol=45, adjCol=c(1,1), 
+          breaks=(-20:20)/20, margins = c(8, 8), cexRow=1.5, cexCol=1.5,font=2)
+
+dev.off()
+}
 ```

### 2022-03-20 manuscript.Rmd
updates to readme

@@
   pdf_document: 
+    highlight: "kate"
     toc: yes
-    toc_depth: '4'
+    toc_depth: 4
   html_notebook:

### 2022-03-20 functions.R

@@ -1,0 +1,442 @@
+###Should list the required packages here
+
+#take a named vector and convert to a data frame with the names turned into 
+#a column
+vector_to_dataframe <- function(vec, new_col = 'V2'){
+  vec <- as.data.frame(vec) 
+  vec[[new_col]] <- row.names(vec)
+  rownames(vec) <- NULL
+  return(vec)
+}
+
+#select and order columns of a data frame using a vector of column names
+select_columns_by_list <- function(dat, vec){
+  dat <- dat[, colnames(dat) %in% vec]
+  dat <- dat[, match(vec, colnames(dat))]
+  return(dat)
+}
+
+#set first letter to capital
+firstup <- function(x) {
+  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
+  x
+}
+
+#take an input, and calculate the cumulative counts of rows sorted by the 
+#target column
+cumulativeCounts <- function(dists, smooth = T, target_column = 'max_dist'){
+  groups <- unique(dists$group)
+  for(i in groups){
+    dat <- dists %>% filter(group == i)
+    dat <- dat %>% mutate(count = 1) %>%
+      arrange(-.data[[target_column]]) %>% group_by(group) %>%
+      mutate(cumulativeCount = cumsum(count)) %>% ungroup() %>%
+      group_by(group, .data[[target_column]]) %>% 
+      summarise(cumulative_prop = max(cumulativeCount)/ nrow(dat))
+    if(smooth){
+      dat <- as.data.frame(spline(x = dat[[target_column]],
+                                  y =  dat$cumulative_prop))
+    }
+    dat <- dat %>% ungroup() %>% mutate(group = i)
+    if(exists('combinedDat')){
+      combinedDat <- combinedDat %>% bind_rows(dat)
+    }else{
+      combinedDat <- dat
+    }
+  }
+  return(combinedDat)
+}
+
+#plot the output of the cumulative counts and run a K-S test
+cumulativeDistribution <- function(dat, run.ks.test = F, 
+                                   alternative_pred, 
+                                   alternative_pc, 
+                                   alternative = 'two.sided', 
+                                   show.legend = F, 
+                                   target_column = 'max_dist'){
[…]
